Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2015-05-03 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Ch5

For piloted ignition tests, an electric spark ignitor is positioned at the top of vertical samples and over the center of horizontal samples. Samples are typically 100~mm squares. Combustion products and dilution air are extracted through the hood and exhaust duct by a high temperature fan. The flow rate can be adjusted between 0.01 and 0.03~m$^3$/s. The volumetric flow rate is kept constant during testing. The sample is mounted on a load cell to determine mass loss rate during a test, and smoke obscuration is measured using a laser light source. The gas flow rate in the exhaust duct is calculated from the pressure drop across and temperature at an orifice plate in the duct. Finally, the concentrations of oxygen, carbon dioxide, carbon monoxide, were measured using appropriate instruments. Heat release rate is calculated from the gas concentration and mass flow measurements.

For this study samples were cut from painted gypsum wallboard panels that were prepared with the panels used in the compartment experiments. The experiments were conducted with the sample in the horizontal position.  Measurements made with the cone calorimeter, in addition to heat release rate per unit area of the samples, included using a series of experiments at different heat fluxes to determine a ignition time and surface temperature pairs.

@@

The radiant flooring panel test method is known as ASTM~E648, Standard Method of Test for Critical Radiant Flux of Floor-Covering Systems Using a Radiant Heat Energy Source~\cite{ASTM_E648}. A 1~m long test specimen is mounted horizontally beneath a pre-mixed air-gas-fueled radiant panel. The panel is inclined at 30~$^{\circ}$, and under standard operating conditions provide a heat flux on the test specimen varying from 10~kW/m$^2$ at the point closest to the panel to 1~kW/m$^2$ at the farthest point. For these experiments a higher heat flux profile was required, so the gas air increased    comback   to  Figure~\ref{FRP_HF_Calibration}

A pilot flame is used to ignite the test sample. The objective of the test is to determine the critical radiant flux for a given sample.  The critical radiant flux corresponds to the flux at the point of maximum flame propagation, or the point on the sample where the flame spread stopped. For this study the samples examined were cut from painted gypsum wallboard panels that were prepared with the panels used in the compartment experiments.

\begin{figure}[p]
	\centering
	\includegraphics[width=\textwidth]{../Figures/FRP_HF_Calibration} \\
	\caption[Photographs of the fire effects on the instrumented gypsum wallboard exposed to replicate polyurethane foam fires]{Photographs of the fire effects on the instrumented gypsum wallboard exposed to replicate polyurethane foam fires with the burner positioned 0.0~m from the wall.}
	\label{FRP_HF_Calibration}
\end{figure}

@@

Fire Dynamics Simulator~\cite{FDS_Users_Guide} is a computational fluid dynamics (CFD) model developed and maintained by NIST that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires. Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells. The cells are contained together within a larger volume known as the computational domain. The CFD model computes the density, velocity, temperature, pressure, and gas concentrations in each cell. Based on the laws of conservation of mass, momentum, and energy, the model tracks the generation and movement of fire gases. One of the most important aspects of FDS is that it is mathematically verified~\cite{FDS_Verification_Guide}and validated against fire test data~\cite{FDS_Validation_Guide}to ensure that it is accurate and provides the expected results, given appropriate input data. A complete description of the FDS model is provided in the FDS Technical Reference Guide~\cite{FDS_Tech_Guide}.

Smokeview is a software tool designed to visualize simulation results from FDS~\cite{Smokeview_Users_Guide}. Smoke-view visualizes smoke and other attributes of the fire simulation by showing tracer particle flow and two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data, including temperature and flow vectors showing flow direction and magnitude. Smokeview has the ability to visualize fire and smoke by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS. Smokeview can also visualize data at particular snapshots in time using 2D or 3D contours of data that show temperature, flow direction, and flow magnitude.

Input data was collected from the experimental results presented in Chapters XXs and from literature sources in order to simulate the experiments a fire using FDS or any other fire model. For the simulation results presented in this document, information was obtained from two primary sources

@@

@@

>>>>

For piloted ignition tests, an electric spark ignitor is positioned at the top of vertical samples and over the center of horizontal samples.  Combustion products and dilution air are extracted through the hood and exhaust duct by a high temperature fan. The flow rate can be adjusted between 0.01 and 0.03~m$^3$/s. The volumetric flow rate is kept constant during testing. The sample is mounted on a load cell to determine mass loss rate during a test, and smoke obscuration is measured using a laser light source. The gas flow rate in the exhaust duct is calculated from the pressure drop across and temperature at an orifice plate in the duct. Finally, the concentrations of oxygen, carbon dioxide, carbon monoxide, were measured using appropriate instruments. Heat release rate is calculated from the gas concentration and mass flow measurements.

For this study samples were cut from painted gypsum wallboard panels. The samples were 100~mm square, and 12.7~mm thick. The experiments were conducted with the samples in the horizontal position.  The key measurements and observations made with the cone calorimeter were heat release rate per unit area of the samples and time to ignition. These experiments were conducted with heat fluxes of 25, 35, 50, 70 and 75 kW/m$^2$. Several other cone calorimeter experiments were conducted with a single type K thermocouple, installed in a slit in the painted paper on the gypsum wallboard samples.  This was the same manner of thermocouple installation used in the instrumented gypsum board wall experiments.  These experiments were conducted at heat fluxes of 25, 30, and 35 kW/m$^2$ to determine ignition time and surface temperature pairs.

@@

The radiant flooring panel test method is known as ASTM~E648, Standard Method of Test for Critical Radiant Flux of Floor-Covering Systems Using a Radiant Heat Energy Source~\cite{ASTM_E648}. A 1~m long test specimen is mounted horizontally beneath a pre-mixed air-gas-fueled radiant panel. The panel is inclined at 30~$^{\circ}$, and under standard operating conditions provide a heat flux on the test specimen varying from 10~kW/m$^2$ at the measurement point, centered 100~mm from the edge of the sample closest to the panel to 1~kW/m$^2$ at the farthest measurement point 900~mm from the edge of the sample closest to the panel. For these experiments a higher heat flux profile was required, so the gas air increased to provide an incident flux of approximately 21~kW/m$^2$ to 2~~kW/m$^2$. Figure~\ref{FRP_HF_Calibration}

A pilot flame from a mm long line burner is used to ignite the test sample near the edge of the sample closet to the panel. The objective of the test is to determine the critical radiant flux for a given sample.  The critical radiant flux corresponds to the flux at the point of maximum flame propagation, or the point on the sample where the flame spread stopped. For this study the samples examined were cut from painted gypsum wallboard panels that were prepared from the supply of panels that were used in the compartment experiments.

\begin{figure}[p]
	\centering
	\includegraphics[width=\textwidth]{../Figures/FRP_HF_Calibration} \\
	\caption[Incident heat flux profile in the radiant flooring panel test apparatus]{Incident heat flux profile along the centerline of the sample in the radiant flooring panel test apparatus}
	\label{FRP_HF_Calibration}
\end{figure}

@@

Fire Dynamics Simulator~\cite{FDS_Users_Guide} is a computational fluid dynamics (CFD) model developed and maintained by NIST that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires. Within a CFD model, the computational domain is divided into small three-dimensional rectangular control volumes or computational cells. The CFD model computes the density, velocity, temperature, pressure, and gas concentrations in each cell. Based on the laws of conservation of mass, momentum, and energy, the model tracks the generation and movement of fire gases. One of the most important aspects of FDS is that it is mathematically verified~\cite{FDS_Verification_Guide}and validated against fire test data~\cite{FDS_Validation_Guide}to ensure that it is accurate and provides the expected results, given appropriate input data. A complete description of the FDS model is provided in the FDS Technical Reference Guide~\cite{FDS_Tech_Guide}.

Smokeview is a software tool designed to visualize simulation results from FDS~\cite{Smokeview_Users_Guide}. Smoke-view visualizes smoke and other attributes of the fire simulation by showing tracer particle flow and two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data, including temperature and flow vectors showing flow direction and magnitude. Smokeview has the ability to visualize fire and smoke by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS. Smokeview can also visualize data at particular snapshots in time using 2D or 3D contours of data that show temperature, flow direction, and flow magnitude.  Smokeview also has a feature that can represent a fire pattern on a surface based on user supplied threshold criteria.

@@

An FDS model was developed to simulate the instrumented wall experiments with the natural gas burner.  Three different computational cell resolutions were examined to determine which resolution provided the best match to both the data from the instruments as well as the data from the fire patterns generated in the wall experiments.  The three computational grid sizes used were : 80~mm, 40~mm, and 20~mm.

@@

Input data was collected from the experimental results presented in Chapters II and V, and from literature sources in order to simulate the instrumented wall experiments using FDS.

Geometry – the domain of the model was cubic with 2.4~ m on a side.  The resulting number computational cells for each case ranged from 27,000 for the 80~mm case to 1,728,000 for the 20~mm cases.   An isometric view of the model outlined by the domain is shown in Figure     .  The face of the wall is located 1.35~m from the front of the domain.  
The wall was modelled with material properties for gypsum board.
Thermal conductivity = 0.28
Specific Heat = 1.00
Density = 582

Instruments – FDS has devices, which are used to simulate thermocouples and total heat flux gauges.  This devices were positioned  in locations similar to the locations used in the physical experiments, in the plume region and on the wall.   A radiative heat flux gauge was also modelled to provide data at the 0.95 m from the centerline of the burner and 1.0 m above the top of the burner.  
In addition, FDS features devices that can not be implemented such as a device, which can be used to examine the heat release rate per unit length.  In this case, this device was used to determine flame height based on that measure.     
Fire 
The burner was modeled as an obstruction, with the same physical size as the burner.  It was modelled with a steady state heat release rate of 888.89 kW/sq m.   The chosen fuel reaction was `methane’ since the majority <<<check percentage>> of the natural gas used in the laboratory experiments was methane.
